Enhanced Material Point Method to face dynamic problems:\thanks{Funding: This study was funded by Agustín de Betancourt Foundation (grant number 262390106114).}

### 2020-03-14 ListaFiguras.tex
Pdf con fotos

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
% flush right qed marks, e.g. at end of proof
\smartqed

\usepackage[lofdepth,lotdepth]{subfig}
% \usepackage{graphicx}
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

@@

%\usepackage{latexsym}
% etc.

%

% Matrix and vector nodes
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}

% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}

% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}

% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

%
% Insert the name of "your journal" with
\journalname{Computational Particle Mechanics}
%
\begin{document}

% \title{Novel improvements in the Material Point Method to face
%   dynamic problems: 
%   Local-maximum entropy approximation and an explicit predictor-corrector scheme. \thanks{Funding: This
%     study was funded by Agustín de Betancourt Foundation (grant number
%     262390106114).}
% }

\title{Enhanced Material Point Method to face
  dynamic problems:\thanks{Funding: This
    study was funded by Agustín de Betancourt Foundation (grant number
    262390106114).}
}
\subtitle{Local-maximum entropy approximation and an explicit
  predictor-corrector scheme}

\titlerunning{Novel improvements in the Material Point Method to face
  dynamic problems.} % if too long for running head

\author{Miguel Molinos \and
  Pedro Navas \and
  Manuel Pastor \and
  Miguel Martín Stickle
}

%\authorrunning{Short form of author list} % if too long for running head

% \institute{Miguel Molinos \at
%   E.T.S de Ingenieros de Caminos, Canales y Puertos. \\
%   Universidad Politéctnica de Madrid, 28040 Madrid, Spain \\
%   \email{m.molinos@alumnos.upm.es}
%   % \emph{Present address:} of F. Author  %  if needed
%   \and
%   Pedro Navas 
%   \and
%   Manuel Pastor
%   \and
%   Miguel Martín Stickle
% }
% \institute{Miguel Molinos, Pedro Navas and Manuel Pastor \at E.T.S
%   de Ingenieros de Caminos, Canales y Puertos, Universidad Politéctnica de Madrid, 28040 Madrid, Spain, \; \email{m.molinos@outlook.es, pedro.navas@upm.es, manuel.pastor@upm.es}
% }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle
\begin{abstract}
  This document is devoted to describe the suitability of the Local
  \textit{maximum-entropy} (LME) meshfree approximation technique under the framework of the
  Material Point Method for dynamic problems. 
  \keywords{LME \and MPM \and Dynamic problems}
  % \PACS{PACS code1 \and PACS code2 \and more}
  % \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Since the proposal of the Material Point Method (\textbf{MPM}) by
Sulsky {\it  et al.} (1994)\cite{Sulsky1994} as a generalization to
solids of the fluid implicit particle (FLIP) method
\cite{Brackbill1986}. It popularity has increased due to its ability
to deal with large strain regime without suffer mesh distortion
inaccuracies. However, this method suffers other kind of
instabilities, such those when material points crossing cell
boundaries. This give rise to the development of other interpolation
techniques to overcome this limitation such as the generalized
interpolation material point method (GIMP) Bardenhagen \& Kober
(2004)\cite{Bardenhagen2004}, the dual
domain material point method (DDMP) Zhang {\it et al.}
(2011)\cite{Zhang2011a}, the B-Spline MPM Tran {\it et al.}
(2019)\cite{Tran2019a} or the Conservative Taylor
Least Squares reconstruction Wobbes {\it et al.}
(2018)\cite{E_Wobbes_2018}. This techniques are devoted to mitigate the
``grid crossing'' error.
Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using techniques \cite{Tran2019e}. These numerical
inaccuracies also known as wiggles are due to the time discretization
technique. A simple approach to face those spurious noises is to add a
damping source to the equilibrium equations. To overcome this
limitation, many researchers has proposed alternative time integration
schemes which reduce them by filtering high-frequencies. Some of these
attempts are the implicit GIMP (iGIMP) Charlton {\it et al.}
(2017)\nocite{Charlton2017} or the generalised-$\alpha$ scheme for the
MPM proposed by Tran \& Solowski (2019)\nocite{Tran2019e}.

The aim of this paper is to mitigate this spurious oscillations by the
employ of a suitable combination of the maximum-entropy (or local
\textit{max-ent}) $\gamma$ family shape functions, and the proposal of a explicit 
predictor-corrector scheme. Local \textit{max-ent} shape functions was
first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite elements (\textbf{FE}) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness of the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions \cite{Arroyo2006}. Furthermore, as
it is not based in a voxel discretization as the GIMP, it is no prone to suffer
voxel domains overlap, which can introduce severe inaccuracies. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)\cite{Navas2018a} and Li {\it et al.} (2012)\cite{Li2012} for
Optimal Transportation Meshfree (\textbf{OTM}) method. And more
recently under the MPM framework by Wobbes {\it et
  al.}(2020)\cite{Wobbes2020} but limiting the $\gamma$ parameter to
zero. On the other hand, the time integration scheme here proposed is an accommodation of
the Newmark a-form ($\gamma = 0.5$ and $\beta = 0$)\footnote{Notice
that these $\gamma$ and $\beta$ parameters does not have the same
meaning for the LME shape functions.} which is the central difference
explicit (CD). We have choose this method among other suitable
alternatives as Wilson {\it et al.} (1972)\nocite{Wilson1972} or Chung
\& Hulbert (1993)\nocite{Geranlized_alpha_1993} because it simplicity
and it good performance dealing with solids dynamics problems like those
presented in \cite{Navas2018a}.

The article is organized as follows. Section \ref{sec:Notation}
briefly reviews the notation here employed. Next, Section \ref{sec:derivation-mpm}
is devoted to present briefly the governing equations of the elastic
problem, the variational formulation and the Galerkin procedure. In Section \ref{sec:epc-algor-mpm} an explicit predictor-corrector
time integration scheme for the MPM is proposed. Section
\ref{sec:local-max-ent} briefly reviews the local \textit{max-ent}
basis functions here employed. In Section \ref{sec:dynamic-benchmark}
a comparison between both methods MPM and OTM employing \textit{max-ent} shape
function is performed as well a parametric study for the $\gamma$
parameter. Both techniques will face to solve a classic dynamic
benchmark proposed by Dyka \& Ingel (1995)\cite{Dyka1995}. Finally,
conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

\section{Brief note concerning the notation}
\label{sec:Notation}
In what follows, we will adopt the following convention. All the
physical variables involved in this problem are collected in Table
\ref{tab:notation-table}.

\begin{table}
  \centering
  \caption{Physical variables involved in the problem}
  \label{tab:notation-table}
  \begin{tabular}{c c c}
    $\rho$ & Density field & Scalar \\
    $\vect{a}$ & Acceleration field & First order tensor \\
    $\vect{v}$ & Velocity field & First order tensor \\
    $\vect{u}$ & Displacement field & Second order tensor \\
    $\tens{\sigma}$ & Cauchy stress tensor & Second order tensor \\
    $\tens{\varepsilon}$ & Cauchy strain tensor & Second order tensor \\
    $\tens{D}$ & Constitutive tensor & Fourth order tensor\\
  \end{tabular}
\end{table}

Three kind of subscript or superscript are used within paper. The
subscript $\qed_p$ is used to define a particle variable. While the
subscript $\qed_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\qed^{\psi}$ involves a virtual
magnitude. For the operators, the convention is : $\dot{\qed}$ and
$\ddot{\qed}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $(\cdot)$ and $(\colon)$ means the first
and second contraction of a tensor, $\Div{\qed}$
denotes the divergence operator, and finally $\Grad{\qed}$ and
$\GradS{\qed}$ denotes the gradient and its symmetric part. Einstain
subscripts convention is adopted so repeated index means addition.

\section{Derivation of the MPM}
\label{sec:derivation-mpm}
The aim of this section is to provide an overview of
the standard explicit MPM algorithm \cite{Sulsky1994}. Therefore, it
is structured in the following shape: we will first give a
introduction to the governing equations \ref{sec:governing-equations},
later the variational statement of the problem will be presented
\ref{sec:variational-formulation}, and finally a discretization
procedure thorough Galerkin is performed \ref{sec:Galerkin-procedure}.

The method has three main steps: (i) a variational recovery process,
where particle data is projected to the grid nodes, (ii) an Eulerian
step, where balance of momentum equation is expressed as a nodal
equilibrium equation thorough a FEM-like procedure, and finally (iii) a Lagrangian advection of the particles.

\subsection{Governing equations }
\label{sec:governing-equations}
In the MPM approach a continuum is considered. Let define a
$\varOmega$ domain occupied by an elastic body, and $\partial
\varOmega$ the boundary of the domain, where $\partial
\varOmega = \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n =
\emptyset$. Note that $\varGamma_d$ is reserved for Dirichlet boundary
conditions, and  $\varGamma_n$ for the Neumann. In a inifinitesimal
strain context, the behaviour of the continuum can be described by the
following set of equations:
(i) the balance of momentum equation
\begin{equation}
  \label{eq:Balance-momentum}
\rho \vect{a} = \Div{\tens{\sigma}} + \rho \vect{b},
\end{equation}
(ii) the compatibility equation with ensure that the velocity
field is compatible with the Cauchy strain field,
\begin{equation}
  \label{eq:Compatibility-equation}
  \dot{\tens{\varepsilon}} = \GradS{\vect{v}},
\end{equation}
(iii) the constitutive equation, which relates the rate of Cauchy strain
tensor with the rate of Cauchy stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\dot{\tens{\sigma}} = \tens{D} \colon \dot{\tens{\varepsilon}},
\end{equation}
and finally (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

\subsection{Variational formulation}
\label{sec:variational-formulation}
To write the variational statement of the problem, let us define a
virtual velocity field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{v}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{v}^{\psi} \in
  \mathcal{H}^1 \mid \vect{v}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{v}^{\psi}} < \infty\ \quad\text{and}\quad \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary value problem of elasticity is the function $\vect{v} \in
\mathcal{H}^1_0$ such that, for $\vect{v}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \Deriv{\vec{v}}{t}\ - \vec{b} \right) \cdot \vec{v}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{v}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
    \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) is the balance of momentum formulated in
its weak form.

\subsection{Galerkin procedure}
\label{sec:Galerkin-procedure}
In order to arrive to a finite set of equations, in contrast with the
FEM, in the MPM a double discretization procedure is performed as we will
describe here below. First, the continuum domain
$\Omega$ is discretized with a finite sum of material points (in the
following particles), each one
represent a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point $\vec{x}_p$ is defined at the centroid
of each $\Omega_p$, figure \ref{fig:MPM-discretization}.
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Discretization_MPM}
  }
  \caption{MPM discretization.}
  \label{fig:MPM-discretization}
\end{figure}

Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$, but also the
virtual field $\vect{v}^{\psi}_{p}$. So employing the definition of the
material integral, where we recover the Riemann integral definition as
an addition  of a finite set of points. Therefore individual terms in
\eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \vec{a} \cdot \vect{v}^{\psi}} =
    \dot{\vec{v}}_{p} \cdot \vect{v}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{v}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{v}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{v}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{v}^{\psi}} = \vec{t}^s_{p} \cdot \vect{v}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Here is
where the second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated. Where $N_n$ is the number of grid
nodes. This mesh is employed as a support to compute gradients and divergences.

Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces}, 
and \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
and approximating the real and virtual velocity  field of the particle
$p$ as
\begin{equation}
  \label{eq:Approx-velocity-mesh}
  \vec{v}_{p} = N_I(\vec{x}_p) \vect{v}_I
  \quad \text{and} \quad \vec{v}^{\psi}_{p} = N_I(\vec{x}_p) \vect{v}_I^{\psi},
\end{equation}
\begin{equation}
  \label{eq:Approx-gradvelocity-mesh}
  \Grad{\vect{v}}_p = \vect{v}_I N_I(\vec{x}_p)
  \quad \text{and} \quad \vec{v}^{\psi}_{p} = \sum^{Nn}_I N_I(\vec{x}_p) \vect{v}_I^{\psi},
\end{equation}
we reach to the particle balance of forces of the continuum,
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$, the nodal mass matrix $\tens{m}_{IJ}$ is,
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} =
  \sum^{Np}_{p} N_{Ip} m_p N_{Jp}.
\end{equation}
To improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_{IJ}^{lumped}$.
Later, internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \sum^{N_p}_{p} \tens{\sigma}_{p} \cdot \Grad{N_{Ip}} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = \sum^{N_p}_{p} N_{Ip} \vec{b}_{p} m_p  + \sum^{N_p}_{p}
  N_{Ip} \vec{t}^s_{p} m_p h^{-1} 
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor is updated employing the rate of stress tensor $\dot{ \tens{\varepsilon}}_{p}$ used to update the
strain tensor is as follows \eqref{eq:IncrStrainPoint}.
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2} \sum^{Nn}_I \left[\Grad{N_{Ip}}\ \otimes \vec{v}_{I} + \vec{v}_{I} \otimes
    \Grad{N_{Ip}}\ \right].
\end{equation}
Next, imposing $\frac{D \rho}{D t} = 0$, we ensures the mass
conservation and update the density field.
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{tra} \left( \dot{\tens{\varepsilon}} \right)
\end{equation}

Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order
temporal integration scheme is required. Therefore, time is
discretized in to a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation it is solved, the values in the nodes are
interpolated back in to the particles and each particle it is advected
to the new position,
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = \sum^{N_n}_{I}N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = \sum^{N_n}_{I}N_{Ip}\ \vec{v}_{I}  
\end{equation}
In the MPM literature, this equations
(\ref{eq:particle_balance_forces3}) and (\ref{eq:Updated_Lagrangian}),
are solved with an explicit forward Euler algorithm. The time integration scheme of the MPM has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in Figure \ref{fig:MPM_algorithm}.

\begin{figure}\sidecaption
  \centering
  \resizebox{0.55\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme}
  }
  \caption{MPM standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}

Other authors have proposed many others time integration alternatives
like \cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the
first publication on the MPM \cite{Sulsky1994}, the nodal acceleration
was employed to update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k}
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} point out, this algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I} \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}\frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},
\end{equation}
Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for the MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I} N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + 
  \sum^{N_n}_{I}N_{Ip}^{k} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right]
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vect{a}_p^{k+1} = \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

This scheme has prof to damps out the higher frequency noises
\cite{Tran2019e}. But it can present the same numerical instabilities
as in \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v} when
nodal masses become infinitesimal.

\section{Explicit predictor-corrector scheme for MPM.}
\label{sec:epc-algor-mpm}

In this section, an explicit predictor-corrector time integration
scheme is presented. It is based in the Newmark a-form 
$\gamma = 0.5$ and $\beta = 0$ which is the central difference
explicit. This method is devoted to solve a system of equations of type
\begin{equation*}
  \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}.
\end{equation*}

As the MPM has a nodal stage, it is possible to apply this methods
successfully in the MPM framework as was proved by
\cite{Tran2019e}. Taking the predictor definition from the classic
literature \cite{Hughes2000} and calculating nodal velocity, and
updating particles position employing nodal values of velocity and
acceleration. Therefore we get the following expression for the
\textbf{predictor} stage for both velocity and displacement.
\begin{equation}
  \label{eq:5}
  \vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
  \underbrace{\sum^{N_n}_{I}N_{Ip}^k\ \vec{v}_{I}^{k}}_{\vec{v}_p^{k}} + \frac{1}{2}\Delta
  t^2\ \underbrace{\sum^{N_n}_{I}\frac{N_{Ip}^k\ \vec{f}_{I}^{k}}{\tens{m}_I^k}}_{\vec{a}_p^{k}}
\end{equation}
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{v}_I^{k+1} = \vec{v}_I^n + (1 - \gamma)\ \Delta t\ \vec{a}_I^k
  \quad , \quad \vec{a}_I^k = \frac{\vec{f}_{I}^{k}}{\tens{m}_I^k}
\end{equation}

Consequently, the \textbf{corrector} stage for the nodal velocity is in the
following way
\begin{equation}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{k+1} + \gamma\ \Delta t\
  \vec{a}_I^{k+1} \quad , \quad  \vec{a}_I^{k+1} = \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}} 
\end{equation}

Notice that the predictor of the particle displacement is
computed using information only from the predictor of the
previous step. However, velocities are computed using half information
of the past corrector and half of the past predictor. Therefore here we share
similarities with the \textbf{leapfrog integration} which updates the
position at full time step, but updates the velocity at half time
steps. Notice also that with this approach the calculation of nodal
momentum vales in not required, and particles velocities is just a
post-process. So during the lagrangian step only the particles
position is truly computed. This minimizes the information
transference between the particles and nodes, and in consequence
minimizes the associated error to this process.

Due to its simplicity allows be implemented with minor modifications
over a standard forward Euler. It is summarized in shape of
pseudo-algorithm \ref{alg:PCE-algorithm}.

>>>>

\documentclass{article}

\usepackage[lofdepth,lotdepth]{subfig}

%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

@@

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Discretization_MPM}
  }
  \caption{Discretization used in the material point method.}
  \label{fig:MPM-discretization}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{0.4\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme}
  }
  \caption{Classic material point method algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
  \centering
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[Piecewise linear (Q4).]{
    \includegraphics[width=0.23\columnwidth]{Figures/MPM_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/MPM_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/MPM_Shape_Fun_dy}
    \label{fig:MPM_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[Local \textit{max-ent} $\gamma=17$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_17_3_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_17_3_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_17_3_Shape_Fun_dy}
    \label{fig:LME_17.3_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[Uniform GIMP (uGIMP).]{
    \includegraphics[width=0.23\columnwidth]{Figures/GIMP_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/GIMP_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/GIMP_Shape_Fun_dy}
    \label{fig:GIMP_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subfloat[Local \textit{max-ent} $\gamma=10$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_10_0_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_10_0_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_10_0_Shape_Fun_dy}
    \label{fig:LME_10.0_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
  \subfloat[Local \textit{max-ent} $\gamma=7$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_7_0_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_7_0_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_7_0_Shape_Fun_dy}
    \label{fig:LME_7.0_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
  \subfloat[Local \textit{max-ent} $\gamma=5$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_5_0_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_5_0_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_5_0_Shape_Fun_dy}
    \label{fig:LME_5.0_Shape_Fun}
  }
  \caption{Comparison of the local \textit{max-ent} shape function for
    different values of $\gamma = \beta h^2$, the piecewise linear shape function
    and the uniform GIMP shape function. The picture shows the nodal value
    of each shape function and its derivatives evaluated in a material
    point located in the center of the domain.}
  \label{fig:LME_MPM}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
